Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling

### 2021-04-08 manuscript/manuscript.Rmd
bunch of edits

---
title: Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - categorical data
  - neural networks
  - embeddings
  - attention
  - severity modeling
  - National Flood Insurance Program 
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
    \newcommand{\norm}[1]{\| #1 \|}
---

@@

On the other hand, several open questions about the use of embeddings within actuarial work remain, which we aim to address in this study. First, hyperparameter settings for embeddings, such as the dimensions of the embedding layer and the use of regularization techniques such as dropout or normalization, that achieve optimal predictive performance has not yet been studied in detail in the actuarial literature. In this work, we aim to study how embeddings using different settings perform in the context of a large-scale predictive modeling problem, and give guidance on the process that can be followed to determine this in other problems. Although neural network have been shown to achieve excellent predictive accuracy on actuarial tasks, many actuaries still prefer to use GLM models for pricing tasks, thus, the issue of whether transferring embeddings to GLM models can achieve better performance is considered in this paper. <!-- Traditional actuarial techniques such as credibility theory have been used to work with some types of categorical data, but no study has been --> <!-- performed whether this can be applied within embeddings; here, we investigate whether performance is enhanced by applying credibility theory on embeddings 
relating to categorical variables with many labels. Whereas embedding layers are usually considered in the context of categorical data,the 
option exists to quantize numerical data and model it using embeddings, however, the results of doing so have not been investigated. --> Finally, in the past several years, a new type of neural network architecture based on attention [@Vaswani2017] has been successfully used on embeddings in the field of natural language processing and we incorporate attention based models into our predictive modeling example to demonstrate the use of these models for actuarial purposes.

In this work, we utilize the recently released National Flood Insurance Program (NFIP) dataset [@FederalEmergencyManagementAgency2019] which provides exposure information for policies written under the NFIP, as well as the claims data relating to these exposures. We refer the reader to Appendix \ref{EDA} for an exploratory analysis of this dataset.

The rest of this manuscript is organized as follows. Section \ref{lit_review} reviews recent applications of embeddings in the actuarial literature. Section \ref{defns} provides the notation used in the paper and defines GLMs, neural networks and related modeling concepts, including embeddings and attention. In Section \ref{modeling}, we provide initial models for the NFIP dataset and consider the influence of hyperparameter choices on the results of the neural network model. That section also considers how successfully the embeddings used in the neural network model can be transferred to a GLM model of the same data and how learned embeddings can be interpreted. We focus on attention based models in Section \ref{attention_models}. Finally, Section \ref{Conclusions} provides a discussion of the results of this paper and considers avenues for future research.

@@

Contemporaneous with that work is the DeepTriangle model of @Kuo2019, which applied recurrent neural networks to the problem of Incurred but not Reported (IBNR) loss reserving, to model jointly the paid and incurred losses in the Schedule P dataset. Embedding layers were used to capture the effect of differences in reserving delays and loss ratios for each company in the Schedule P dataset. Evaluating the results of the DeepTriangle method showed that the out of sample performance of the model (tested against the lower triangles in the Schedule P dataset) exceeded that of traditional IBNR reserving techniques.

@@

The predictor variables that we consider here are comprised of two types: continuous variables, taking on numerical values and represented by the matrix $X_{num}$ with $J^{num}$ columns, and categorical variables, which take on discrete values indicating one of several possible categories, represented by the matrix $X_{cat}$ with $J^{cat}$ columns, such that $J^{num} + J^{cat} = J$.

@@

A categorical variable $X_j, j \in J_{cat}$ takes as its value only one of a finite number of labels. Let the set of labels be $\mathcal{P}^j = \{p^j_1, p^j_2,\dots, p^j_{n_{\mathcal{P}^j}} \}$, where $n_\mathcal{P}^j = |\mathcal{P}^j|$ is the cardinality or number of levels, in $\mathcal{P}^j$. One-hot encoding maps each value $x_{i,j}$ of $X_j$ to $n_\mathcal{P}^j$ indicator variables, which take a value of $1$ if the label of $x_{i,j}$ corresponds to the level of the indicator variable, and $0$ otherwise. An example of one-hot encoding is shown in Table \@ref(tab:onehot).

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is made available by the OpenFEMA initiative of the Federal Emergency Management Agency (FEMA)[^1]. Two datasets are made available by OpenFEMA: A policies dataset with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to data from 2000 to 2019 and further subsample 100K data points to make experiments feasible on a CPU. The dataset includes a rich variety of variables, from occupancy type to flood zone. For our models, we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, which we list in Table \@ref(tab:variables)

@@

We note that the focus of our contribution is on applications of categorical embeddings in predictive modeling, rather than on fine-tuning models in order to beat or establish a benchmark. To this end, we do not perform systematic hyperparameter tuning in our experiments, and the performance metrics are provided to show that reasonable models utilizing categorical embeddings can be developed (note in the next section, the models achieve significantly better performance). However, as we describe the modeling procedure and results, we provide commentary on recommended practices in practical applications, especially with respect to the neural network models.

@@

```{r occ-type, echo=FALSE, fig.cap="Learned embeddings of occupancy type visualized on a number line.", out.width = "80%", fig.align="center"}
knitr::include_graphics("images/occupancy_type.png")
```

```{r flood-zone, echo = FALSE, fig.cap="Learned embeddings of flood zone by prefix.", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone.png")
```

@@

While t-SNE is a powerful visualization technique, there are a few points the modeler should be mindful of in practice. We refer the reader to @wattenberg2016how for a more thorough discussion, but stress here that the resulting plot can be very sensitive to the perplexity hyperparameter. In addition, for our use case, which is to explore the learned embeddings, we have fewer data points than most other applications in the literature. Accordingly, our reasonable range of hyperparameter values need to be adjusted down. In Figures \@ref(fig:flood-zone-tsne), we display the t-SNE of the learned embeddings for `flood_zone` in Model 3 with various perplexity settings: 2, 3, 5, and 10; the learning rate and the number of steps are fixed at 100 and 10,000, respectively, for all plots. In the first three plots, we can see the semblance of clusters forming. In particular, flood zone D appears in its own cluster, which implies it is far away from other levels in the embedding space. When the perplexity is set to 10 (and beyond), however, we start to lose the geometric properties.

@@

Whereas Model 5 makes a simple modification to the structure of Model 4, the TabTransformer model is slightly more complex. With respect to similarities between Model 5 and 6, both use embeddings with a standard 8 dimensions for each categorical covariate. In terms of differences, Model 6 uses a Transformer layer in place of self-attention and a positional encoding of a single dimension, i.e. $d_col = 1$. The TabTransformer architecture is shown in Figure \@ref(fig:nn-3).

@@

To investigate the predictive value added by the contextual embeddings, a simple Gamma GLM was fit to these 10 000 observations, with the goal of predicting the claim amount based only on either the embeddings or the contextual embeddings. As shown in Table \@ref(tab:emd-vs-cembd), the performance of the latter model, as indicated by the RMSE and MAE metrics is significantly better, in other words, the contextual embeddings that incorporate information across all of the categorical variables are significantly better predictors of claim severity compared with the embeddings.

>>>>

---
title: Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevinykuo@gmail.com 
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants and University of the Witwatersrand
    email: ron@ronaldrichman.co.za
abstract: |
  Categorical data are ubiquitous in insurance modeling. In this paper, we explore in depth how these data can be
  processed with categorical embeddings in the context of claim severity modeling. We develop several models that range
  in complexity from simple neural networks to state-of-the-art attention based architectures and calibrate categorical
  embeddings using these models. We illustrate the utility of these learned embeddings as pretrained features in a
  generalized linear model. We discuss methods for visualizing and interpreting learned embeddings. Finally, we explore
  how attention based models can contextually augment embeddings, leading to enhanced predictive performance.
keywords:
  - categorical data
  - neural networks
  - embeddings
  - attention
  - severity modeling
  - National Flood Insurance Program 
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
    \newcommand{\norm}[1]{\| #1 \|}
---

@@

On the other hand, several open questions about the use of embeddings within actuarial work remain, which we aim to address in this study. First, hyperparameter settings for embeddings, such as the dimensions of the embedding layer and the use of regularization techniques such as dropout or normalization, that achieve optimal predictive performance has not yet been studied in detail in the actuarial literature. In this work, we aim to study how embeddings using different settings perform in the context of a large-scale predictive modeling problem, and give guidance on the process that can be followed to determine this in other problems. Although neural network have been shown to achieve excellent predictive accuracy on actuarial tasks, many actuaries still prefer to use GLM models for pricing tasks, thus, the issue of whether transferring embeddings to GLM models can achieve better performance is considered in this paper. Finally, in the past several years, a new type of neural network architecture based on attention [@Vaswani2017] has been successfully used on embeddings in the field of natural language processing and we incorporate attention based models into our predictive modeling example to demonstrate the use of these models for actuarial purposes.

In this work, we utilize the recently released National Flood Insurance Program (NFIP) dataset [@FederalEmergencyManagementAgency2019] which provides exposure information for policies written under the NFIP, as well as the claims data relating to these exposures.

The rest of this manuscript is organized as follows. Section \ref{lit_review} reviews recent applications of embeddings in the actuarial literature. Section \ref{defns} provides the notation used in the paper and defines GLMs, neural networks and related modeling concepts, including embeddings and attention. In Section \ref{modeling}, we provide initial models for the NFIP dataset and consider how successfully the embeddings used in the neural network model can be transferred to a GLM model of the same data and how learned embeddings can be interpreted. We focus on attention based models in Section \ref{attention_models}. Finally, Section \ref{Conclusions} provides a discussion of the results of this paper and considers avenues for future research.

@@

Contemporaneous with that work is the DeepTriangle model of @Kuo2019, which applied recurrent neural networks to the problem of Incurred but not Reported (IBNR) loss reserving, to model jointly the paid and incurred losses in the NAIC Schedule P dataset. Embedding layers were used to capture the effect of differences in reserving delays and loss ratios for each company in the Schedule P dataset.

@@

The predictor variables that we consider here are comprised of two types: continuous variables, taking on numerical values and represented by the matrix $X_{num}$ with $|J^{num}|$ columns, and categorical variables, which take on discrete values indicating one of several possible categories, represented by the matrix $X_{cat}$ with $|J^{cat}|$ columns, such that $|J^{num}| + |J^{cat}| = J$.

@@

A categorical variable $X_j, j \in J^{cat}$ takes as its value only one of a finite number of labels. Let the set of labels be $\mathcal{P}^j = \{p^j_1, p^j_2,\dots, p^j_{n_{\mathcal{P}^j}} \}$, where $n_\mathcal{P}^j = |\mathcal{P}^j|$ is the cardinality or number of levels, in $\mathcal{P}^j$. One-hot encoding maps each value $x_{i,j}$ of $X_j$ to $n_\mathcal{P}^j$ indicator variables, which take a value of $1$ if the label of $x_{i,j}$ corresponds to the level of the indicator variable, and $0$ otherwise. An example of one-hot encoding is shown in Table \@ref(tab:onehot).

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid on the property coverage of the policy. The data we use comes from the NFIP and is made available by the OpenFEMA initiative of the Federal Emergency Management Agency (FEMA)[^1]. Two datasets are made available by OpenFEMA: A policies dataset with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to data from 2000 to 2019 and further subsample 100K data points to make experiments feasible on a CPU. The dataset includes a rich variety of variables, from occupancy type to flood zone, and a brief exploratory data analysis can be found in Appendix \ref{EDA}. For our models, we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, which we list in Table \@ref(tab:variables)

@@

We note that the focus of our contribution is on applications of categorical embeddings in predictive modeling, rather than on fine-tuning models in order to beat or establish a benchmark. To this end, we do not perform systematic hyperparameter tuning in our experiments, and the performance metrics are provided to show that reasonable models utilizing categorical embeddings can be developed. However, as we describe the modeling procedure and results, we provide commentary on recommended practices in practical applications, especially with respect to the neural network models.

@@

```{r occ-type, echo=FALSE, fig.cap="Learned embeddings of occupancy type visualized on a number line.", out.width = "80%", fig.align="center"}
knitr::include_graphics("images/occupancy_type_embed.png")
```

```{r flood-zone, echo = FALSE, fig.cap="Learned embeddings of flood zone by prefix.", out.width = "60%", fig.align = "center"}
knitr::include_graphics("images/flood_zone_embed.png")
```

@@

While t-SNE is a powerful visualization technique, there are a few points the modeler should be mindful of in practice. We refer the reader to @wattenberg2016how for a more thorough discussion, but stress here that the resulting plot can be very sensitive to the perplexity hyperparameter. In addition, for our use case, which is to explore the learned embeddings, we have fewer data points than most other applications in the literature. Accordingly, our reasonable range of hyperparameter values need to be adjusted down. In Figures \@ref(fig:flood-zone-tsne), we display the t-SNE of the learned embeddings for `flood_zone` in Model 3 with various perplexity settings: 2, 3, 5, and 10; the learning rate and the number of steps are fixed at 100 and 10,000, respectively, for all plots. In the first three plots, we can see the semblance of clusters forming. In particular, flood zone D appears in its own cluster, which implies it is far away from other levels in the embedding space, *as seen by the model*. When the perplexity is set to 10 (and beyond), however, we start to lose the geometric properties.

@@

Whereas Model 5 makes a simple modification to the structure of Model 4, the TabTransformer model is slightly more complex. With respect to similarities between Model 5 and 6, both use embeddings with a standard 8 dimensions for each categorical covariate. In terms of differences, Model 6 uses a Transformer layer in place of self-attention and a positional encoding of a single dimension, i.e. $d_{col} = 1$. The TabTransformer architecture is shown in Figure \@ref(fig:nn-3).

@@

To investigate the predictive value added by the contextual embeddings, a simple Gamma GLM was fit to these 10 000 observations, with the goal of predicting the claim amount based only on either the embeddings or the contextual embeddings. As shown in Table \@ref(tab:emd-vs-cembd), the performance of the latter model, as indicated by the RMSE and MAE metrics is better, in other words, the contextual embeddings that incorporate information across all of the categorical variables are better predictors of claim severity compared with the embeddings.
